---
title: "Analyzing textual data"
output: html_document
---

In this lab we will work through an example of text mining. First we will download the complete text of Romeo and Juliet:
```{r}
randj <- readLines("http://www.gutenberg.org/cache/epub/1112/pg1112.txt")
```

Let's remove the header information from the file and inspect beginning of the play.
```{r}
randj <- randj[-(1:65)]
head(randj, 20)
```
**Exercise**. How many lines are there in the play? Don't count empty lines.




#Zipf's law
In the first part of this lab we will create a histogram of the word frequency distribution i.e. how many words appear only once, how many words appear twice etc. 

**Step 1** Notice that play is provided as a list of lines. Split each line into a vector so that each entry in the vector contains only one word (Hint recall the ``str.split`` function.)

```{r, include = FALSE}
randj <- strsplit(randj, split = " ")
```

**Step 2** Concatenate all of the vectors into a single long vector.

```{r, include = FALSE}
randj <- do.call(c, randj)
randj <- tolower(randj)
```

**Step 3** Use table to find out how many times each word appears in the vector.
```{r, include = FALSE}
counts <- table(randj)
freqs <- table(table(randj))
```


**Step 4** Use table a second time to find out how many words appear only once, how many appear twice etc.

**Step 5** Plot your results as a histogram. You should get something that looks like this>

```{r, fig.keep='last', echo = FALSE}
plotdd <- as.data.frame(freqs)
names(plotdd) <- c("appearances", "freq")
plotdd$appearances <- as.numeric(as.character(plotdd$appearances))
library(ggplot2)
ggplot() +
  geom_bar(aes(x = as.vector(counts)), binwidth = 1) +
  xlim(1,50)
```


# Making a word cloud.
For this part we will be using the package ```wordcloud```. Let's load it now.
```{r, warning = FALSE}
library(wordcloud)
```

The function we want is called ```wordcloud``` too. Have a look at the helpfile. 

```{r}
?wordcloud
```

**Exercise** Use the ```wordcloud``` function to make a wordcloud for 100 most common words in Romeo and Juliet. Make sure to the empty string "" is not in your list of words! It should look something like this:

```{r, echo = FALSE, warning = FALSE}
counts <- sort(counts, decreasing = TRUE)[-1]
wordcloud(names(counts)[1:100], as.vector(counts)[1:100],scale=c(8,0.3),
          colors=brewer.pal(8, "Dark2"))
```

Notice that there are are many words that are not all that interesting. Fortunately some people have already collected a list of these very common words. They call them stop words:

```{r, warning = FALSE}
#install.packages("tm")
library(tm)
stopwords(kind = "en")
```

**Exercise** Remake your word cloud excluding the these very common words.

```{r, include = FALSE, warning = FALSE}
counts_small <- counts[!names(counts) %in% stopwords(kind = "en")]
wordcloud(names(counts_small), as.vector(counts_small), max.words = 100,scale=c(8,0.3),
          colors=brewer.pal(8, "Dark2"))
```

